Longitudinal Principal Manifold Estimation

### 2024-04-03 paper/main.tex
update adni volume figure references

To understand the impacts of these discrepancies on the estimated volume of the regions, the volume of the hippocampus was approximated from the raw data and the estimates obtained by the PME and LPME algorithms at each time point. The volume estimation approach used here relies on counting the voxels contained within the boundary defined by the PME- or LPME-estimated embedding map, with each voxel having a known volume. The interior identification approach developed by \cite{mengPrincipalManifoldEstimation2021} is used to determine whether each voxel may be considered inside the surface of the structure or not. The code used to implement this approach is available at \texttt{https://github.com/rjzielinski/lpme-project}. The volume estimates can be found in Figure \ref{fig:lhipp_volumes}. Here, we see that volume estimates differ substantially depending on whether PME, LPME, or the raw data were used, with the LPME estimate resulting in the lowest volume computations.

@@

Considering the same individual shown in Figures \ref{fig:adni_result} and \ref{fig:adni_cross_section}, results of fitting the PME and LPME algorithms to the surface of the left thalamus are shown in Figures \ref{fig:adni_lthal_result} and \ref{fig:lthal_cross_sections}. While the segmentation estimates of the thalamus do not show the same sort of deviations that are clearly visible for the hippocampus, the cross section plots shown in Figure \ref{fig:lthal_cross_sections} indicate that both the PME and LPME algorithms show a closer fit to the more regularly shaped thalamus surface than was achieved for the hippocampus. This is reflected in the smaller discrepancies between volume estimates, as seen in Figure \ref{fig:lthal_volumes}, though visible variations between time points remain.

>>>>

To understand the impacts of these discrepancies on the estimated volume of the regions, the volume of the hippocampus was approximated from the raw data and the estimates obtained by the PME and LPME algorithms at each time point. The volume estimation approach used here relies on counting the voxels contained within the boundary defined by the PME- or LPME-estimated embedding map, with each voxel having a known volume. The interior identification approach developed by \cite{mengPrincipalManifoldEstimation2021} is used to determine whether each voxel may be considered inside the surface of the structure or not. The code used to implement this approach is available at \texttt{https://github.com/rjzielinski/lpme-project}. The volume estimates can be found in Figure \ref{fig:lhipp_volume_comparison}. Here, we see that volume estimates differ substantially depending on whether PME, LPME, or the raw data were used, with the LPME estimate resulting in the lowest volume computations.

@@

Considering the same individual shown in Figures \ref{fig:adni_result} and \ref{fig:adni_cross_section}, results of fitting the PME and LPME algorithms to the surface of the left thalamus are shown in Figures \ref{fig:adni_lthal_result} and \ref{fig:lthal_cross_sections}. While the segmentation estimates of the thalamus do not show the same sort of deviations that are clearly visible for the hippocampus, the cross section plots shown in Figure \ref{fig:lthal_cross_sections} indicate that both the PME and LPME algorithms show a closer fit to the more regularly shaped thalamus surface than was achieved for the hippocampus. This is reflected in the smaller discrepancies between volume estimates, as seen in Figure \ref{fig:lthal_volume_comparison}, though visible variations between time points remain.
